---
created by: Shi Yunn
---
PA1
==============================

Setting

```{r}
echo = TRUE  # Always make code visible
options(scipen = 1)  # Turn off scientific notations for numbers
```


Loading and preprocessing the data

```{r}
setwd("C:/Users/Asus/Desktop/Data Science Coursera/Reproducible Research/")
unzip("C:./repdata-data-activity.zip")
activity <- read.csv("activity.csv")
library(dplyr)
```

Removing NA steps from data frame

```{r}
activity1 <- filter(activity, !is.na(steps))
```

Grouping the dataset by date

```{r}
activity_date <- group_by(activity1, date)
```

Calculate the total steps per day

```{r}
activity_steps <- aggregate(steps ~ date, activity1, sum)
colnames(activity_steps) <- c("date", "steps")
head(activity_steps)
```

Plot the histrogram to illustate the number of steps per day

```{r}
library(ggplot2)
ggplot(activity_steps, aes(x=steps)) + 
  geom_histogram(binwidth = 1000)+
  labs(title= " The number of steps taken per day (NA removed",
       x = "Num. of steps per day",
       y = " Num. of times in day") 
```

To calculate the mean and median of the steps taken per day

```{r}
mean(activity_steps$steps)
median(activity_steps$steps)
```

The mean is 10766 and median is 10765

To find what is the average daily activity pattern

```{r}
activity_pattern <- aggregate(activity1$steps,
                              by = list(interval= activity1$interval),
                              FUN=mean, na.rm=TRUE)
colnames(activity_pattern) <- c("interval", "steps")
head(activity_pattern)
```

Plot a line graph to track the av. daily activity pattern

```{r}
ggplot(activity_pattern, aes(x=interval, y=steps)) + 
  geom_line(size=1) + 
  labs (title= "Av. daily activity pattern(NA removed)", x="Interval", y="Num. of steps")
```

To find out which 5-min interval, on average across all the days in the dataset, contains the maximum number of steps?

```{r}
max_int <- activity_pattern[which.max(activity_pattern$steps),]
```

The answer is 835th interval with a max. of 206 steps

Calculate the total number of missing values in the dataset (i.e. the total number of rows with NAs)

```{r}
na_steps <- sum(is.na(activity$steps))
```

the total number of missing values is 2304

Devise a strategy for filling in all of the missing values in the dataset. In this case, I'm using mean of the steps to replace NA

```{r}
mean_na <- rep(mean(activity$steps, na.rm=TRUE), times=length(which(is.na(activity$steps))))

activity[which(is.na(activity$steps)), "steps"] <- mean_na

head(activity)
```

Make a histogram of the total number of steps taken each day

```{r}
activity_fill <- aggregate(activity$steps, by=list(activity$date), FUN=sum)
colnames(activity_fill) <- c("date", "total")

ggplot(activity_fill, aes(x=total)) + 
  geom_histogram(binwidth = 1000)+
  labs(title= " The number of steps taken per day (NA replaced)",
       x = "Num. of steps per day",
       y = " Num. of times in day")
```

Calculate and report the mean and median total number of steps taken per day.

```{r}
mean(activity_fill$total)
median(activity_fill$total)
```

The mean and median are 10766. Mean remains unchanged, but median has increased. After filling up the NA values, mean and median become equal.

Create a new factor variable in the dataset with two levels - "weekday" and "weekend"

```{r}
library(lubridate)
```

Lubridate is an R package that makes it easier to work with dates and times.

```{r}
activity <- mutate(activity, weekend=factor(weekdays(ymd(date)) %in% c("Saturday", "Sunday")))
```

group by weekend and interval

```{r,echo=TRUE}
grp_activity <- group_by(activity, weekend, interval)
grp_activity_mean <- group_by(grp_activity, steps=mean(steps))

```

Make a panel plot containing a time series plot (i.e. type = "l") of the 5-minute interval (x-axis) to show the differences between weekday and weekend

```{r, echo=TRUE}
wday <- ggplot(data=subset(grp_activity_mean, weekend==FALSE), aes(x=interval, y=steps)) + geom_line() + 
  ggtitle("Weekday")

wend <- ggplot(data=subset(grp_activity_mean, weekend==TRUE), aes(x=interval, y=steps)) + geom_line() + 
  ggtitle("Weekend")
```

to compare both plots

```{r}
library(gridExtra)
grid.arrange(wday, wend, ncol=1)
```

Generate PA1_template.md and PA1_template.html files

```{r}
install.packages('knitr', dependencies = TRUE)
library(knitr)
knit2html("PA1_template.Rmd")
```


